
%%%%%%%%% MASTER -- compiles the 4 sections

\documentclass[11pt,letterpaper]{article}

\usepackage{graphicx}
\usepackage{verbatim}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pagestyle{plain}                                                      %%
%%%%%%%%%% EXACT 1in MARGINS %%%%%%%                                   %%
\setlength{\textwidth}{6.5in}     %%                                   %%
\setlength{\oddsidemargin}{0in}   %% (It is recommended that you       %%
\setlength{\evensidemargin}{0in}  %%  not change these parameters,     %%
\setlength{\textheight}{8.5in}    %%  at the risk of having your       %%
\setlength{\topmargin}{0in}       %%  proposal dismissed on the basis  %%
\setlength{\headheight}{0in}      %%  of incorrect formatting!!!)      %%
\setlength{\headsep}{0in}         %%                                   %%
\setlength{\footskip}{.5in}       %%                                   %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                   %%
\newcommand{\required}[1]{\section*{\hfil #1\hfil}}                    %%
\renewcommand{\refname}{\hfil References Cited\hfil}                   %%
\bibliographystyle{plain}                                              %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%PUT YOUR MACROS HERE

\date{December 14th, 2011}
\title{Evolutionary Computation\\
in Muscle Action Simulations
}

\author{Colby Blair \\
	Computer Science Undergraduate \\
	University of Idaho Computer Science Department
}

\begin{document}
\maketitle

\begin{abstract}
In biomechanics, there are many interactions between the body and brain. Some
interactions can be quantified as electrical signals to motor nerves of muscles.
The result is a particular musclular reaction due to an electrical signal. The
input is an electrical nerve signal, and the output is a measured velocity at
a joint. The overall result to this sytem of inputs and output is actions like 
walking.

In this project, a Genetic Algorithm (GA) is used to replace the Simulated Annealing (SA) optimization
that Dr. Craig McGowan uses for his research in biomechanics. His research takes EMG signals as inputs,
and tries to predict joint velocities as outputs. This report shows that the GA is much slower, but 
maintains better solutions when it is set up with the right parameters. This report will also recommend
further research opportunities for better results and performance.
\end{abstract}

\thispagestyle{empty}


\pagebreak

\thispagestyle{empty}
\tableofcontents
\listoffigures

\pagebreak

\setcounter{page}{1}
\include{reportIntro}
\include{reportDesc}
\include{reportResults}
\include{reportConc}
\include{reportRefs}
%\include{reportBio}



\end{document}
